Designing Large-Scale Imaging Surveys for a Retrospective Relative Photometric Calibration

### 2012-06-04 paper/referee_response.txt

excellent suggestions, improve the paper. Respond inline

	--Sec 2.2: "current" generation of large survey imagers would be more
	  accurate than "next" generation.

Done.

	--Sec 3.1: "We We"

Done.

	--Sec 5.1: please include some description of the density and/or total
	  number of sources used, either here or in Sec. 2.2.  Also insert "16
	  deg^2" before "patch of sky".

We explicitly put the source density into the last sentence of 2.1 and we re-point the reader to the parameter table in the last sentence prior to Section 5.1. We included the solid angle of the total patch, as requested.

	--suggestion for going a bit deeper in Sec 5.1: some surveys may find
	  it difficult to rotate their hardware randomly with every shot.
	  Could you do a survey E which has random pointing centers but no
	  rotation?

The referee is absolutely right that there are cameras that cannot be rotated with respect to the sky. That is one of the reasons why we designed Survey C without rotation. Since Survey C calibrates just as well as Survey D, we conclude that rotation is not necessary for good self-calibration. We agree that a Survey E would confirm that, but we believe that Survey C already makes this point well. We added a few sentences to the discussion of Survey C and D about this point in Sections 5.1 and 5.2.

	--suggestion for going a bit deeper in Sec 5.2 (and/or Sec 4): you
	  show the residuals in the instrument response, but more relevant
	  might be the residuals in the final survey.  For Survey A (and to a
	  large extent B), the residuals in the final survey are just as bad
	  as in the instrument response.  But in Surveys C and D, they get beat
	  down because the calibration error on a given piece of sky changes
	  from observation to observation.  It would be instructive to add a
	  performance metric which includes this effect.

In Addition to these changes, we also made the following small changes:
	- MeasuredCatalog.true_invvar - multiplied by flat-field one time too many??
	- Generate epsilon now per measurement and not per source

Neither made significant changes to content of papers

### 2012-06-04 paper/self_calibration_I.tex
response to all apart from one referee comment

\begin{abstract}
We study the self-calibration -- the determination of a complex system
response from science data alone -- for precise photometric catalogs
from wide-field imaging surveys. We create an artificial sky of sources
and synthetically observe it under four basic survey strategies,
creating an end-to-end simulation of an imaging survey for each. These
catalog-level simulations include realistic measurement uncertainties
and a complex focal-plane dependence of the instrument response. In the
self-calibration step, we simultaneously fit for all the star fluxes
and the parameters of a position-dependent flat-field.  For realism,
we deliberately fit with a wrong noise model and a flat-field
functional basis that does not include the model that generated the
synthetic data.  We demonstrate that with a favorable survey strategy,
a complex (but smooth) instrument response can be precisely
self-calibrated. We show that returning the same sources to very
different focal plane positions is the key property of any survey
strategy designed for accurate retrospective calibration. The results
of this work suggest the following advice for those considering the
design of large-scale imaging surveys: Do not use a regular tiling of
the sky; instead return the same sources to very different focal plane
positions.
\end{abstract}

@@

\subsection{Synthetic Sky}
We generate a representative synthetic sky based on realistic object densities in the AB magnitude range $m_\text{min} = 17$ to $m_\text{max} = 22$~mag$_\text{AB}$, with these limits chosen to be consistent with the saturation and a $10\upsigma{}$ limits of deep, space-based, near-infrared survey. Sources are generated with random coordinates (uniformly distributed within the sky region being investigated) and with random magnitudes $m$ distributed according to
\begin{eqnarray}
\log_{10} \frac{\dd N}{\dd m \, \dd \Omega} = a + b\,m + c\,m^2 \label{eqn:power_law} \quad , 
\end{eqnarray}
where $\dfrac{\dd N}{\dd m \, \dd  \Omega}$  is the density of sources $N$ per unit magnitude $m$ and per unit solid angle $\Omega$, and $a$, $b$ and $c$ are model parameters. Even though our simulations make no distinction between galaxies and stars, the values of the parameters are found from fitting the Y-band galaxy populations reported in \citet{win11} only. These parameters were found to be $a = -13.05$, $b = 1.25$ and $c = -0.02$. The source magnitudes $m$ are related to the source fluxes $s$ simply by: $m = 22.5 - 2.5\log_{10}(s)$, where the 22.5 puts the fluxes in units of nanomaggies (nmgy). To limit the data, we only select the brightest sources within the survey area, up to a source density per unit solid angle $d$, for the self-calibration procedure.

@@

\subsubsection{Measured Count Rates}
To generate photometric catalogs, the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position-dependent instrument response model $f_\true$ and a measurement noise model. For a measurement $i$, the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$ -- which is a function of focal plane position $\vec{x_i}$ -- and the source's true flux $s_{k,\true}$ ,
\begin{eqnarray*}
c_i = f_{\true}(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} + e_{i} \quad ,
\end{eqnarray*}
where $\vec{q}_\true$ are the parameters defining the \textit{true} instrument response (see Section \ref{sec:instrument_response_model}), and $e_i$ is a noise contribution drawn from the Normal Distribution $N(e|0,{\sigma_\true}^2)$.

@@

\section{Simple Survey Strategies}
\label{sec:simple_surveys}
In this paper, we consider four simple but very different survey strategies. We use the end-to-end catalog simulation and self-calibration chain, introduced in Sections \ref{sec:survey_simulations} and \ref{sec:self_cal}, to evaluate the performance of the self-calibration procedure with these strategies.

\subsection{Survey Description}
We label the four survey strategies -- which all cover the same 64~deg$^2$ patch of sky -- with the letters A to D. These survey strategies are summarized in Table \ref{tab:simple_surveys} and are shown in Figure \ref{fig:simple_surveys}. Strategy A is the simplest strategy; the target field is regularly tiled with small overlaps between adjacent pointings ($\sim 5$~percent and $\sim 8$~percent in the camera pointing directions $\alpha$ and $\beta$ respectively). The pointings in the 12 passes over the same field are exactly aligned. Survey B is the same as A, but with each pass over the target field the orientation of the telescope is rotated by $30^\circ$. Survey C is more complex. The first pass over the field is the same as in Survey A, with $12 \times 12$ pointings. In the next pass, one of the pointings in the $\alpha$ direction is removed and one is added in the $\beta$ and the pointings are respaced uniformly, so the resultant pointing grid is $13 \times 11$. In the third pass over the field this change is reversed and the field is measured on a $11 \times 13$ grid. These three passes are then repeated four times. The pointing positions in Survey D are quasi-random: the pointings are the same as with Survey A, but each has a random offset within [-0.35,0.35)~deg applied in both the $\alpha$ and $\beta$ directions. By fixing the pointings within these $0.7~\text{deg} \times 0.7~\text{deg}$ boxes, we ensure that the quasi random strategy has a uniform coverage of the field; that is, the fluctuations in coverage are less than they would be in a Poisson (totally random) process. The orientations of the pointings in Survey D are completely random.

@@

The Survey A and B catalogs also give interesting results. With Survey A, the self-calibration procedure did not reach a solution. This -- in itself -- is an important result, as the naive tiling of the sky currently being considered for many large surveys will result in catalogs that cannot be retrospectively self-calibrated. Minor changes to survey strategies, such as those in Survey C, can drastically improve the performance of the self-calibration procedure. Without these changes, vast amounts of calibration information is simply lost. We can also draw conclusions from the results obtained with Survey B: here the self-calibration procedure converges close to the correct solutions, albeit not as close as the Surveys C and D. Here, it is the quality of the fit over the focal plane that is interesting. From the residual plot against the best-in-basis instrument response $f_\basis(\vec{x_j} | \vec{q_\basis})$ (Fig \ref{fig:B_result} bottom right), we can see that the quality of the fit is rotationally symmetric, with the self-calibration procedure able to fit the instrument response well at the edge of the focal plane, but poorly at the center.

@@

In this paper, we have considered the design of survey strategies for calibrating the large-scale, relative photometric response of an instrument, but we feel there are many more calibration parameters that can be constrained from this method. For example, the optical distortion of an instrument or the point-spread-function variation could also be constrained with such a method, although it would need to be explored whether the properties of a survey strategy that makes it good for one calibration, are also suitable for other calibration aspects; ultimately a trade-off would need to be performed between the different calibration concerns and the other goals of the survey, such as sky coverage, variability monitoring and so-on.

\acknowledgments
\section{Acknowledgments}
It is a pleasure to thank Rob Fergus (NYU), Jason Kalirai (STScI),
and Sam Roweis (deceased) for valuable conversations.

@@

\begin{thebibliography}{}
\bibitem[Padmanabhan et al. (2008)]{pad08}{Padmanabhan, N. et al., An Improved Photometric Calibration of the Sloan Digital Sky Survey Imaging Data, 2008, \apj~674, p1217.}
\bibitem[Windhorst et al. (2011)]{win11}{Windhorst, R.~A. et al., The Hubble Space Telescope Wide Field Camera 3 Early Release Science Data: Panchromatic Faint Object Counts for 0.2-2 $\upmu$m Wavelength, 2011, \apjs~193, p27.}
\bibitem[Wittman et al. (2011)]{wit11}{Wittman, D., Ryan, R. and Thorman, P., Ubercalibration of the Deep Lens Survey, 2011, ArXiv e-prints, 1111.2058}
\end{thebibliography}
\clearpage

>>>>

\begin{abstract}
We study the self-calibration---the determination of a complex system
response from science data alone---for precise photometric catalogs
from wide-field imaging surveys. We create an artificial sky of sources
and synthetically observe it under four basic survey strategies,
creating an end-to-end simulation of an imaging survey for each. These
catalog-level simulations include realistic measurement uncertainties
and a complex focal-plane dependence of the instrument response. In the
self-calibration step, we simultaneously fit for all the star fluxes
and the parameters of a position-dependent flat-field.  For realism,
we deliberately fit with a wrong noise model and a flat-field
functional basis that does not include the model that generated the
synthetic data.  We demonstrate that with a favorable survey strategy,
a complex (but smooth) instrument response can be precisely
self-calibrated. We show that returning the same sources to very
different focal plane positions is the key property of any survey
strategy designed for accurate retrospective calibration. The results
of this work suggest the following advice for those considering the
design of large-scale imaging surveys: Do not use a regular tiling of
the sky; instead return the same sources to very different focal plane
positions.
\end{abstract}

@@

\subsection{Synthetic Sky}
We generate a representative synthetic sky based on realistic object densities in the AB magnitude range $m_\text{min} = 17$ to $m_\text{max} = 22$~mag$_\text{AB}$, with these limits chosen to be consistent with the saturation and a $10\upsigma{}$ limits of deep, space-based, near-infrared survey. Sources are generated with random coordinates (uniformly distributed within the sky region being investigated) and with random magnitudes $m$ distributed according to
\begin{eqnarray}
\log_{10} \frac{\dd N}{\dd m \, \dd \Omega} = a + b\,m + c\,m^2 \label{eqn:power_law} \quad , 
\end{eqnarray}
where $\dfrac{\dd N}{\dd m \, \dd  \Omega}$  is the density of sources $N$ per unit magnitude $m$ and per unit solid angle $\Omega$, and $a$, $b$ and $c$ are model parameters. Even though our simulations make no distinction between galaxies and stars, the values of the parameters are found from fitting the Y-band galaxy populations reported in \citet{win11} only. These parameters were found to be $a = -13.05$, $b = 1.25$ and $c = -0.02$. The source magnitudes $m$ are related to the source fluxes $s$ simply by: $m = 22.5 - 2.5\log_{10}(s)$, where the 22.5 puts the fluxes in units of nanomaggies (nmgy). To limit the data, we only select the brightest sources within the survey area, up to a source density $d=100$~deg$^{-2}$, for the self-calibration procedure.

@@

\subsubsection{Measured Count Rates}
To generate photometric catalogs, the true source fluxes $s_\true$ are converted into measured count rates $c$ with an complex, position-dependent instrument response model $f_\true$ and a measurement noise model. For a measurement $i$, the count rate $c_i$ recorded from a source $k$ depends on the \textit{true} instrument response $f_{\true}(\vec{x_i} | \vec{q}_\true)$---which is a function of focal plane position $\vec{x_i}$---and the source's true flux $s_{k,\true}$ ,
\begin{eqnarray*}
c_i = f_{\true}(\vec{x_i} | \vec{q}_\true) \, s_{k, \true} + e_{i} \quad ,
\end{eqnarray*}
where $\vec{q}_\true$ are the parameters defining the \textit{true} instrument response (see Section \ref{sec:instrument_response_model}), and $e_i$ is a noise contribution drawn from the Normal Distribution $N(e|0,{\sigma_\true}^2)$.

@@

\section{Simple Survey Strategies}
\label{sec:simple_surveys}
In this paper, we consider four simple but very different survey strategies. We use the end-to-end catalog simulation and self-calibration chain, introduced in Sections \ref{sec:survey_simulations} and \ref{sec:self_cal}, to evaluate the performance of the self-calibration procedure with these strategies. The parameters of these simulations are summarized in Table \ref{tab:parameters}.

\subsection{Survey Description}
We label the four survey strategies---which all cover the same 64~deg$^2$ patch of sky---with the letters A to D. These survey strategies are summarized in Table \ref{tab:simple_surveys} and are shown in Figure \ref{fig:simple_surveys}. Strategy A is the simplest strategy; the target field is regularly tiled with small overlaps between adjacent pointings ($\sim 5$~percent and $\sim 8$~percent in the camera pointing directions $\alpha$ and $\beta$ respectively). The pointings in the 12 passes over the same field are exactly aligned. Survey B is the same as A, but with each pass over the target field the orientation of the telescope is rotated by $30^\circ$.

Survey C is more complex. The first pass over the field is the same as in Survey A, with $12 \times 12$ pointings. In the next pass, one of the pointings in the $\alpha$ direction is removed and one is added in the $\beta$ and the pointings are respaced uniformly, so the resultant pointing grid is $13 \times 11$. In the third pass over the field this change is reversed and the field is measured on a $11 \times 13$ grid. These three passes are then repeated four times.

The pointing positions in Survey D are quasi-random: the pointings are the same as with Survey A, but each has a random offset within [-0.35,0.35)~deg applied in both the $\alpha$ and $\beta$ directions. By fixing the pointings within these $0.7~\text{deg} \times 0.7~\text{deg}$ boxes, we ensure that the quasi random strategy has a uniform coverage of the field; that is, the fluctuations in coverage are less than they would be in a Poisson (totally random) process. The orientations of the pointings in Survey D are completely random.

Surveys A and C can be executed without rotating the camera with respect to the celestial coordinates. Surveys B and D can only be executed with systems that permit rotations of the camera.

@@

Surveys C and D perform comparably well, although they are very different in their configuration. Two substantial advantages of Survey C over Survey D are: In Survey C the camera would not need to be rotated relative to the sky between pointings; indeed many survey imaging cameras have fixed celestial orientations. Survey D requires generating quasi-random numbers, whereas Survey C has a completely deterministic set of pointing centers. This is a conceptual advantage to Survey C, but possibly also an operational advantage.

The Survey A and B catalogs also give interesting results. With Survey A, the self-calibration procedure did not reach a solution. This---in itself---is an important result, as the naive tiling of the sky currently being considered for many large surveys will result in catalogs that cannot be retrospectively self-calibrated. Minor changes to survey strategies, such as those in Survey C, can drastically improve the performance of the self-calibration procedure. Without these changes, vast amounts of calibration information is simply lost. We can also draw conclusions from the results obtained with Survey B: here the self-calibration procedure converges close to the correct solutions, albeit not as close as the Surveys C and D. Here, it is the quality of the fit over the focal plane that is interesting. From the residual plot against the best-in-basis instrument response $f_\basis(\vec{x_j} | \vec{q_\basis})$ (Fig \ref{fig:B_result} bottom right), we can see that the quality of the fit is rotationally symmetric, with the self-calibration procedure able to fit the instrument response well at the edge of the focal plane, but poorly at the center.

@@

In this paper, we have considered the design of survey strategies for calibrating the large-scale, relative photometric response of an instrument, but there are many more calibration parameters that can be constrained from this method. For example, the optical distortion of an instrument or the point-spread-function variation could also be constrained with such a method, although it would need to be explored whether the properties of a survey strategy that makes it good for one calibration, are also suitable for other calibration aspects; ultimately a trade-off would need to be performed between the different calibration concerns and the other goals of the survey, such as sky coverage, variability monitoring and so-on.

\acknowledgments
It is a pleasure to thank Rob Fergus (NYU), Jason Kalirai (STScI),
and Sam Roweis (deceased) for valuable conversations.

@@

\begin{thebibliography}{}
\bibitem[Padmanabhan et al.(2008)]{pad08}{Padmanabhan, N. et al., An Improved Photometric Calibration of the Sloan Digital Sky Survey Imaging Data, 2008, \apj~674, p1217.}
\bibitem[Windhorst et al.(2011)]{win11}{Windhorst, R.~A. et al., The Hubble Space Telescope Wide Field Camera 3 Early Release Science Data: Panchromatic Faint Object Counts for 0.2-2 $\upmu$m Wavelength, 2011, \apjs~193, p27.}
\bibitem[Wittman et al.(2011)]{wit11}{Wittman, D., Ryan, R. and Thorman, P., Ubercalibration of the Deep Lens Survey, 2011, ArXiv e-prints, 1111.2058}
\end{thebibliography}
\clearpage
